%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{Symbols}
\label{app:symbols}

This appendix provides an exhaustive list of symbols used throughout this dissertation.
Symbols are listed by chapter and sorted alphabetically. Where applicable, Roman
symbols are listed first, followed by Greek symbols. In cases where symbols are re-used,
they are re-defined under the relevant chapters.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Chapter~\ref{chap:divisionoflabour}: Division Of Labour}
\label{sec:symbols:divisionoflabour}


\begin{description}
	\item[\parbox{\namewidth}{$R$}] The set of tasks that can be performed by individuals of the swarm.
	\item[\parbox{\namewidth}{$\gamma$}] An individual robot in a swarm.
	\item[\parbox{\namewidth}{$\nu$}] A task such that $\nu \in R$
	\item[\parbox{\namewidth}{$r_{\gamma,\nu}$}] A response threshold for of robot $\gamma$ for task $\nu$.
	\item[\parbox{\namewidth}{$z$}] The probability of a robot to switch from search behaviour to rest behaviour.
\end{description}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Chapter~\ref{chap:third}: Nature Inspired Algorithms for Prioritized Foraging}
\label{sec:symbols:foraging}

\begin{description}
\setlength{\itemsep}{-1mm}
	\item[\parbox{\namewidth}{$i$}] The current time step in a swarm robot experiment
	
\item[\parbox{\namewidth}{$state$}] The state a particular robot is in.

\item[\parbox{\namewidth}{$role$}] The role of a particular robot in the honey bee algorithm.

	\item[\parbox{\namewidth}{$i_{state}$}] The current time step of being in consecutive state $state$.
	
	\item[\parbox{\namewidth}{$t_{wait}$}] Maximum time a robot can spend consecutively in wait state.
	
	\item[\parbox{\namewidth}{$t_{ls}$}] Maximum time a robot can spend consecutively in the local search state.
	
	\item[\parbox{\namewidth}{$t_{forage}$}] Maximum time a robot can spend consecutively in the forage state.

	\item[\parbox{\namewidth}{$t_{explore}$}] Maximum time a robot can spend consecutively in the explore state


	\item[\parbox{\namewidth}{$t_{dance}$}] Maximum time a robot can spend consecutively in the recruitment state.
	
	
	\item[\parbox{\namewidth}{$\varsigma$}] The type of an item which can either be prioritized or non-prioritized.

	\item[\parbox{\namewidth}{$\mu_\varsigma$}] The evaluation of the quality of a site of type $\varsigma$.
	
	\item[\parbox{\namewidth}{$\Phi$}] Site quality threshold.
	
	\item[\parbox{\namewidth}{$v$}] A robot's path integration vector

	\item[\parbox{\namewidth}{$\omega$}] A memorized path integration vector representing the location of a site.
	
	\item[\parbox{\namewidth}{$\vartheta$}] An item found by a robot.
	\item[\parbox{\namewidth}{$\xi$}] The site where an item is found.

	\item[\parbox{\namewidth}{$\varrho$}] A random number selected from a uniform distribution such that $\varrho\in(0,1)$.
	
	\item[\parbox{\namewidth}{$\rho$}] The probability that a scout robot becomes a forager robot.
	
\item[\parbox{\namewidth}{$k_j$}] is the value captured on the $j-$th distance sensor of a robot.

\item[\parbox{\namewidth}{$n$}] The number of distance sensors.

\item[\parbox{\namewidth}{$\sigma$}] The direction a robot must take to get to the sink.

\item[\parbox{\namewidth}{$\alpha$}] The probability of listening to the details communicated by the scout robot.		

\item[\parbox{\namewidth}{$f_{max}$}] Maximum time a robot will forage for prioritized items unsuccessfully.

\item[\parbox{\namewidth}{$X$}] The percentage of robots initialized as scout robots.
\end{description}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Chapter~\ref{chap:experiment}: Experimental Setup}
\label{sec:symbols:foraging}


\begin{description}
	\item[\parbox{\namewidth}{$d$}] Desirability of a direction $q$.
	\item[\parbox{\namewidth}{$q$}] A direction a robot can perceive.
	
	\item[\parbox{\namewidth}{$\kappa_q$}] The clarity of a direction.
	\item[\parbox{\namewidth}{$\eta$}] The depth of view of a robot's perception.

	\item[\parbox{\namewidth}{$\iota_q$}] The directness of a direction $q$.

	\item[\parbox{\namewidth}{$f$}] The	field of view of a robot's perception.
		
	\item[\parbox{\namewidth}{$\lambda$}] A ratio between the effect of $\kappa_q$ or $\iota_q$ on $d$.
	
	\item[\parbox{\namewidth}{$\Lambda$}] The height and width of the environment grid.

	\item[\parbox{\namewidth}{$p$}] The density of the items on the grid, also referred to as problem complexity.

	\item[\parbox{\namewidth}{$\overline{p}$}] A list of all considered values for $p$.

	\item[\parbox{\namewidth}{$p_i$}] The $i$th item of $\overline{p}$.

	\item[\parbox{\namewidth}{$p_{min}$}] The smallest considered value for $p$.
	
	
	\item[\parbox{\namewidth}{$r$}] The ratio of prioritized to non-prioritized items.

	\item[\parbox{\namewidth}{$\overline{r}$}] A list of all considered values for $r$.

	\item[\parbox{\namewidth}{$r_i$}] The $i$th item in $\overline{r}$.

	\item[\parbox{\namewidth}{$n_r$}] The number of items in $\overline{r}$	

	\item[\parbox{\namewidth}{$t$}] The time step of an experiment.	

	\item[\parbox{\namewidth}{$\tau$}] The specialization ratio of the robot swarm.

	\item[\parbox{\namewidth}{$\tau(t)$}] The value of $\tau$ at time step $t$.

	\item[\parbox{\namewidth}{$\overline{\tau}$}] A list of all considered values for $\tau(0)$.

	\item[\parbox{\namewidth}{$n_\tau$}] The number of items in $\overline{\tau}$

	\item[\parbox{\namewidth}{$\tau_i(0)$}] The $i$th item in $\overline{\tau}$.

	\item[\parbox{\namewidth}{$Z_{ij}$}] The average efficiency $E_P$ over all experiments with $\tau_i$ and $r_j$.
	
	\item[\parbox{\namewidth}{$Z$}] The matrix of the values $Z_{ij}$ where $i=1,..,n_\tau$ and $j=1,...,n_r$.

	\item[\parbox{\namewidth}{$D_{ij}$}] The average efficiency $E_P$ over all experiments with $\tau_i$ and $\epsilon_j$.
	
	\item[\parbox{\namewidth}{$D$}] The matrix of the values $D_{ij}$ where $i=1,..,n_\tau$ and $j=1,...,n_\epsilon$.
	

	\item[\parbox{\namewidth}{$\epsilon$}] The distribution of prioritized and non-prioritized items on an environment.

	\item[\parbox{\namewidth}{$\overline{\epsilon}$}] A list of all considered environment distributions $\epsilon$.

	\item[\parbox{\namewidth}{$\epsilon_i$}] The $i$th item of $\overline{\epsilon}$.

	\item[\parbox{\namewidth}{$c$}] Density of robots in an environment.

	\item[\parbox{\namewidth}{$\overline{c}$}] A list of all considered values for $c$.
	
	\item[\parbox{\namewidth}{$c_i$}] The $i$th item in $\overline{c}$.

	\item[\parbox{\namewidth}{$c_{min}$}] The smallest value considered value for $c$.
	
	\item[\parbox{\namewidth}{$E_P$}] The percentage of prioritized items foraged over time.

	\item[\parbox{\namewidth}{$E_{NP}$}] The percentage of non-prioritized items foraged over time.

	\item[\parbox{\namewidth}{$I$}] A list where entry $I_j$ is the index yielding the maximum value for $Z_{ij}$.

	\item[\parbox{\namewidth}{$Y$}] A list where entry $Y_j$ is the index yielding the maximum value for $D_{ij}$.

	\item[\parbox{\namewidth}{$F_{r_i}$}] The flexibility of a robot swarm in terms of prioritized item ratio $r_i$.

	\item[\parbox{\namewidth}{$F_r$}] The flexibility of a robot swarm in terms of prioritized item ratio.

	\item[\parbox{\namewidth}{$F_{\epsilon_i}$}] The flexibility of a robot swarm in terms of environment distribution $\epsilon_i$.

	\item[\parbox{\namewidth}{$F_\epsilon$}] The flexibility of a robot swarm in terms of environment distribution.

	\item[\parbox{\namewidth}{$E_{c_i}$}] The average foraging efficiency, $E_P$, with swarm density $c_i$.

	\item[\parbox{\namewidth}{$E_{p_i}$}] The average foraging efficiency, $E_P$, with problem complexity $p_i$.

	\item[\parbox{\namewidth}{$S_c$}] The scalability of a robot swarm in terms of swarm density.	
	
	\item[\parbox{\namewidth}{$S_p$}] The scalability of a robot swarm in terms of problem size.	
	
	\item[\parbox{\namewidth}{$E^t_P$}] Total prioritized items foraged over time $t$.	

	\item[\parbox{\namewidth}{$E^t_{NP}$}] Total non-prioritized items foraged over time $t$.	
	
	\item[\parbox{\namewidth}{$t_{wait}$}] The average time steps spent by each robot in the waiting state.

	\item[\parbox{\namewidth}{$t_{recruitment}$}] The average time steps spent by each robot in the recruitment state. 
	
	

\end{description}
